### 2011-11-22 draft.tex
Wrote paragraphs on the status of intelligence attempts and Dreyfus opinion. Started bibliography collection for the document.

\documentclass{article}
\usepackage{natbib}

\title{Phenomenological Robotics}
\author{Bruno Nery\\
        \texttt{bnery@isr.ist.utl.pt}}

\begin{document}

\maketitle

In the world of Good Old Fashioned Artificial Intelligence (GOFAI), scientists
attempted to fabricate intelligence by doing inference on symbolic
representations of the world. In Brooks \cite{brooks91} approach, this is done
by making agents respond to fixed isolable features of the environment.

Dreyfus \cite{dreyfus07} believes that the basis of human intelligence, coping
with everyday life, cannot be achieved by any of the approaches described above.
Instead, the fact that

\begin{quotation}
\dots embodied beings like us take as input energy from the physical universe
and respond in such a way as to open them to a world organized in terms of their
needs, interests, and bodily capacities, without their minds needing to impose
meaning on a meaningless given \dots nor their brains converting stimulus input
into reflex responses \dots
\end{quotation}

\bibliographystyle{plain}
\bibliography{prsurvey}

\end{document}

### 2011-11-22 fabfile.py
Added fabfile.py.

@@ -1,0 +1,17 @@
+from fabric.api import local
+
+def clean():
+    local("rm draft.aux")
+    local("rm draft.bbl")
+    local("rm draft.blg")
+    local("rm draft.log")
+    local("rm draft.pdf")
+    local("rm fabfile.pyc")
+
+def draft():
+    local("pdflatex draft.tex")
+    local("bibtex draft")
+    local("pdflatex draft.tex")
+    local("pdflatex draft.tex")
+    
+

### 2011-11-25 draft.tex
Wrote on robots having different bodies and different perceptions.

@@
 \documentclass{article}
+\usepackage{graphicx}
 \usepackage{natbib}
@@
 
+% 1.E. Robots have different bodies, different perceptions and different needs
+%      from ours.
+% (Source: Thesis journal, pages 006 to 015, 14/12/2010).
+
+Each robot has a different body, consisting of a set of sensors and actuators. 
+For this reason, a robot experiences the world in a very specific way. Its body
+defines not only what it can do, but also the way in which it sees the world.
+Take the HERB~\cite{srinivasa2009herb}, the iCub~\cite{metta2010icub}, and the
+Pioneer P3-DX robots, for example, depicted in
+Figure~\ref{fig:herb_icub_pioneer}.
+
+\begin{figure}[h]
+\centering
+\includegraphics{figures/herb_icub_pioneer.png}
+\caption{HERB, iCub, and Pioneer P3-DX robots}
+\label{fig:herb_icub_pioneer}
+\end{figure}
+
+The HERB sees the world in \emph{forward mode} through its laser range finder
+and its camera, coupled with its body movement. The iCub, in its
+\emph{paraplegical} version, can only see the world in a left-to-right (and
+right-to-left) manner through its cameras and head/torso movement. It can also
+grasp nearby objects in order to examine them better. Finally, the Pioneer P3-DX
+experiences the world in two different directions at the same time through its
+sonars and its wheels movement.
+
 \bibliographystyle{plain}

### 2011-11-25 fabfile.py
Added check target to fabfile.py.

@@ -14,4 +14,5 @@ def draft():
     local("pdflatex draft.tex")
     local("pdflatex draft.tex")
     
-
+def check():
+    local("evince draft.pdf")

### 2011-12-02 draft.tex
Wrote on the IKEA scenario.

@@
   \dots embodied beings like us take as input energy from the physical universe
-and respond in such a way as to open them to a world organized in terms of their
-needs, interests, and bodily capacities, without their minds needing to impose
-meaning on a meaningless given \dots nor their brains converting stimulus input
-into reflex responses \dots
+  and respond in such a way as to open them to a world organized in terms of
+  their needs, interests, and bodily capacities, without their minds needing to
+  impose meaning on a meaningless given \dots nor their brains converting
+  stimulus input into reflex responses \dots
 \end{quotation}
@@
 
+\section{The IKEA scenario}
+
+\begin{quotation}
+  A hammer is for hammering.
+\end{quotation}
+
+If you have a nail nearby, that might be the case. If you are assembling IKEA
+furniture outdoors, however, a hammer might as well be ``for paper weighting''.
+In the same sense, a ladder is only ``for climbing'' when you have the need to
+reach a high place. Most of the times, a ladder will be
+``for being in the way''.
+
+Also, screwdrivers are, supposedly, ``for driving screws''. But what happens
+when you a have a flat screwdriver and need to drive a Phillips-head screw? It
+is harder to drive it, but have you ever tried driving a flat head screw with a
+Phillips screwdriver? It won't work, and you'll start asking yourself what is
+wrong, and what you can do to fix it.
+
+You will be having a breakdown, as described by Heidegger, and the screwdriver
+will cease to be \emph{ready-to-hand} and become \emph{present-at-hand} until
+you can figure out the problem. But how do we equip a robot with the same
+capabilities so that it can function in a human world?
+
 \bibliographystyle{plain}

### 2011-12-02 draft.tex
Added screwdriver figures to the IKEA scenario.

@@
 \usepackage{natbib}
+\usepackage{subfig}
 
@@
 
+% 4. The IKEA scenario.
+% (Source: Thesis journal, pages 130 to 139, 17/03/2011)
+
 \section{The IKEA scenario}
@@
 Also, screwdrivers are, supposedly, ``for driving screws''. But what happens
-when you a have a flat screwdriver and need to drive a Phillips-head screw? It
-is harder to drive it, but have you ever tried driving a flat head screw with a
-Phillips screwdriver? It won't work, and you'll start asking yourself what is
-wrong, and what you can do to fix it.
+when you a have a flat screwdriver and need to drive a Phillips-head screw
+(Figure~\ref{fig:ikea_flat_phillips})? It is harder to drive it, but have you
+ever tried driving a flat head screw with a Phillips screwdriver
+(Figure~\ref{fig:ikea_phillips_flat})? It won't work, and you'll start asking
+yourself what is wrong, and what you can do to fix it.
 
-You will be having a breakdown, as described by Heidegger, and the screwdriver
-will cease to be \emph{ready-to-hand} and become \emph{present-at-hand} until
-you can figure out the problem. But how do we equip a robot with the same
-capabilities so that it can function in a human world?
+\begin{figure}[h]
+  \centering
+  \subfloat[][Flat screwdriver vs. Phillips-head screw] {
+    \includegraphics{figures/ikea_flat_phillips.png}
+    \label{fig:ikea_flat_phillips}
+  }
+  \qquad
+  \subfloat[][Phillips screwdriver vs. Flat head screw] {
+    \includegraphics{figures/ikea_phillips_flat.png}
+    \label{fig:ikea_phillips_flat}
+  }
+  \caption{The IKEA scenario}
+\end{figure}
+
+You will be having a breakdown, as described by Heidegger \cite{dreyfus07}, and
+the screwdriver will cease to be \emph{ready-to-hand} and become
+\emph{present-at-hand} until you can figure out the problem. But how do we equip
+a robot with the same capabilities so that it can function in a human world?
 

### 2011-12-02 draft.tex
Wrote about the cornerstones of the framework.

@@
 % 4. The IKEA scenario.
-% (Source: Thesis journal, pages 130 to 139, 17/03/2011)
+% (Source: Thesis journal, pages 130 to 139, 17/03/2011).
 
@@
 
-You will be having a breakdown, as described by Heidegger \cite{dreyfus07}, and
+You will be having a breakdown, as described by Heidegger~\cite{dreyfus07}, and
 the screwdriver will cease to be \emph{ready-to-hand} and become
@@
 a robot with the same capabilities so that it can function in a human world?
+
+% Not present in the outline: cornerstones of the framework. This might not be
+% the best place for this part.
+% (Source: Thesis journal, pages 100 to 109, 03/03/2011).
+
+Before a robot is able to handle a breakdown, it needs to be able to function in
+the world seamlessly, fullfiling its needs by executing one action after the
+other, using what is available to it in a given moment. To be able to do that,
+the following set of basic capabilities might be a good place to start:
+
+\begin{itemize}
+\item \textbf{body knowledge:} know its own body, be able to detect its current
+state and how it changes according to its movements;
+\item \textbf{skill acquisition:} know how discover and/or learn a set of
+abilities that allow it to interact with the world;
+\item \textbf{object detection:} know how to single out objects and detect
+properties that allow it to use them in a particular manner.
+\end{itemize}
 

### 2011-12-09 draft.tex
Wrote about the failure of the computation hypothesis on everyday coping.

@@
   stimulus input into reflex responses \dots
+\end{quotation}
+
+% 1.A. "Failure" of the computational hypothesis on everyday coping (being in
+% the world is more basic than thinking).
+% (Source: \cite{dreyfus07}.)
+
+He points out that the pioneers in GOFAI learned, directly or indirectly, from
+the philosophers. Hobbes' claim that reasoning was calculating and Descartes'
+mental representations, among others, are cited as examples. At the same time,
+he states that the critical insights formulated by existentialist philosophers,
+such as Heidegger and Merleau-Ponty, were a sign that their enterprise was
+comdemned to failure.
+
+He criticizes Newell and Simon's claim that the mind could be understood as a
+\emph{physical symbol system}, where streams of neuron pulses would be used as
+symbols to represent the external world. Dreyfus points out that
+
+\begin{quotation}
+  Heidegger's important insight is not that, when we solve problems, we
+  sometimes make use of representational equipment outside our bodies, but that
+  \emph{being-in-the-world} is more basic than \emph{thinking} and solving
+  problems; that it is not representational at all.
 \end{quotation}

### 2011-12-09 draft.tex
Reorganized introduction.

@@
 attempted to fabricate intelligence by doing inference on symbolic
-representations of the world. In Brooks \cite{brooks91} approach, this is done
-by making agents respond to fixed isolable features of the environment.
+representations of the world.
 
-Dreyfus \cite{dreyfus07} believes that the basis of human intelligence, coping
-with everyday life, cannot be achieved by any of the approaches described above.
-Instead, the fact that
+% 1.A. "Failure" of the computational hypothesis on everyday coping (being in
+% the world is more basic than thinking).
+% (Source: \cite{dreyfus07}.)
+
+Dreyfus~\cite{dreyfus07} points out that the pioneers in GOFAI learned, directly
+or indirectly, from the philosophers. Hobbes' claim that reasoning was
+calculating and Descartes' mental representations, among others, are cited as
+examples. At the same time, he states that the critical insights formulated by
+existentialist philosophers, such as Heidegger and Merleau-Ponty, were a sign
+that their enterprise was comdemned to failure. He affirms that
+
+\begin{quotation}
+  Heidegger's important insight is not that, when we solve problems, we
+  sometimes make use of representational equipment outside our bodies, but that
+  \emph{being-in-the-world} is more basic than \emph{thinking} and solving
+  problems; that it is not representational at all.
+\end{quotation}
+
+% 1.?. On Brooks attempt. On Dreyfus believe on the failure of GOFAI and Brooks.
+
+In Brooks' \cite{brooks91} approach, this is done by making agents respond to
+fixed isolable features of the environment. Dreyfus \cite{dreyfus07} believes
+that the basis of human intelligence, coping with everyday life, cannot be
+achieved by any of the approaches described above. Instead, the fact that
 
@@
   stimulus input into reflex responses \dots
-\end{quotation}
-
-% 1.A. "Failure" of the computational hypothesis on everyday coping (being in
-% the world is more basic than thinking).
-% (Source: \cite{dreyfus07}.)
-
-He points out that the pioneers in GOFAI learned, directly or indirectly, from
-the philosophers. Hobbes' claim that reasoning was calculating and Descartes'
-mental representations, among others, are cited as examples. At the same time,
-he states that the critical insights formulated by existentialist philosophers,
-such as Heidegger and Merleau-Ponty, were a sign that their enterprise was
-comdemned to failure.
-
-He criticizes Newell and Simon's claim that the mind could be understood as a
-\emph{physical symbol system}, where streams of neuron pulses would be used as
-symbols to represent the external world. Dreyfus points out that
-
-\begin{quotation}
-  Heidegger's important insight is not that, when we solve problems, we
-  sometimes make use of representational equipment outside our bodies, but that
-  \emph{being-in-the-world} is more basic than \emph{thinking} and solving
-  problems; that it is not representational at all.
 \end{quotation}

### 2011-12-09 draft.tex
Wrote on Merleau-Ponty and relation to E. Gibson. Added dreyfus2002 and gibson2000 to bibliography.

@@
 
+% 2. Relation between phenomenology and alternative psychology theories.
+
+\section{Phenomenology and psychology theories}
+
+% 2.B. Merleau-Ponty skills (refined situations) vs. Eleanor Gibson.
+% (Source: Thesis journal, pages 140 to 149, 29/03/2011).
+% (Source: \cite{gibson2000}.)
+
+Dreyfus~\cite{dreyfus2002} shows that learning and skillful action, deemed as
+basic forms of intelligent behavior, can be explained without recurring to
+mental representations. To accomplish this, he explains the two central concepts
+in Merleau-Ponty's \emph{Phenomenology of Perception}:
+
+\begin{itemize}
+\item \textbf{intentional arc:} is the strong connection between an agent and
+its environment. In that sense, the skills learned by an agent are not kept as
+representations in the mind, but as dispositions to respond to solicitations of
+situations in the environment;
+\item \textbf{maximal grip:} is the tendency of an agent to respond to these
+solicitations in a way that brings the situation status closer to what it
+considers optimal.
+\end{itemize}
+
+Gibson~\cite{gibson2000} propose that the process of learning is one of
+discrimination rather than association, that is, through the refinement of the
+sensorimotor knowledge acquired through the selection of patterns in our
+interaction with the world. According to her, this selection process happens
+based on two criteria:
+
+% (\cite{gibson2000}, pages 154-156).
+
+\begin{itemize}
+\item \textbf{affordance fit:} what increases the fit between the organism and
+its environment, according to its capacities and goals;
+\item \textbf{economy:} what reduces the need for action and perceptual
+information to accomplish a given task. -- optimization
+\end{itemize}
+
+These criteria can be seen as directly related to Merleau-Ponty's
+\emph{intentional arc} and \emph{maximal grip}, respectively.
+
 % 4. The IKEA scenario.

### 2011-12-09 draft.tex
Added list of perceptual learning cases from E. Gibson.

@@
 
+% Not present in the outline: learning process in infancy. This might not be the
+% best place for this part, but it definitely goes together with the
+% cornerstones of the framework below.
+% (Source: Thesis journal, pages 140 to 149, 28/03/2011).
+% (Source: \cite{gibson2000}, pages 134 to 149, chapter 8).
+
+Gibson~\cite{gibson2000} affirms that perceptual learning predates birth. She
+describes a number of cases, going from early perceptual learning to elaborate
+sequences of action in older infants. In these cases, the infant learns:
+
+\begin{itemize}
+\item about oneself;
+\item to control an event;
+\item to perceive an unitary object;
+\item what happens next in an observed event;
+\item to participate in a communicative event;
+\item to differentiate multimodally specified events;
+\item distinctive features of things;
+\item to use an object or an act as means;
+\item about causal relations in observed events.
+\end{itemize}
+
 % Not present in the outline: cornerstones of the framework. This might not be

### 2011-12-16 draft.tex
Wrote about Todes vs. Kuipers. Wrote about Spinozism.

@@
 % (Source: Thesis journal, pages 006 to 015, 14/12/2010).
+% Note: it might make more sense to propose the importance of the robots' bodies
+% after introducing Todes (section 3.A). I'm not sure, however, of how to
+% restructure the text to accomplish that.
 
@@
 
-Gibson~\cite{gibson2000} propose that the process of learning is one of
+Gibson and Pick~\cite{gibson2000} propose that the process of learning is one of
 discrimination rather than association, that is, through the refinement of the
 sensorimotor knowledge acquired through the selection of patterns in our
-interaction with the world. According to her, this selection process happens
+interaction with the world. According to them, this selection process happens
 based on two criteria:
@@
 \emph{intentional arc} and \emph{maximal grip}, respectively.
+
+% 3. In Robotics, Phylosophy and Related Work.
+
+\section{Phenomenology and robotics}
+
+% 3.A. Kuipers vs. Todes.
+% (Source: \cite{todes2001}, page xix).
+% (Source: Thesis journal, pages 42 to 42, 05/01/2011).
+
+Dreyfus, in his introduction to Todes' \emph{Body and World}~\cite{todes2001},
+describes the fundamental role played by the body in structuring the
+spatiotemporal field that defines human experience
+
+\begin{quotation}
+  \dots the structure of the active body produces our unified experience of
+  space and time. Since the body moves forward more effectively than backwards,
+  it opens a \emph{horizontal field} that organizes experience in what can be
+  coped with directly, what can be reached with effort, and what is over the
+  perceptual horizon. Furthermore, the front/back asymmetry of the active body
+  \dots organizes the temporal field.
+\end{quotation}
+
+% Reviewing Dreyfus introduction to Todes' Body and World, I realized that his
+% work contributes a lot more than just the importance of the body and its
+% orientation. According to Dreyfus, he goes further on improving the concepts
+% of intentional arc and maximal grip from Merleau-Ponty. It might be an
+% interesting source of ideas, but I'm not sure they would fit in this text.
+
+Pierce and Kuipers~\cite{pierce1997} developed an algorithm to obtain a robot's
+perception model and its possible actions in an unknown environment. Despite the
+importance given to the robot's body, they make a clear distinction between the
+mind (agent) and the body (robot), seeing the first as learning to control the
+second.
+
+% Not present in the outline: Spinozism and how it counters Pierce and Kuipers.
+% I'm not sure if the way the text below is structured adds much to the point.
+% (Source: \cite{lemmen1997}).
+% ATTENTION: I had access to this thesis via Porfirio Silva, who downloaded it
+% from the author's website in 2002 (no longer available). According to his
+% note, the author asks not to be cited without permission. I found no trace of
+% the author online, with the exception of a paper from 1996 with the email
+% ronaldl@cogs.susx.ac.uk. Chapters 6 and 7 might be worth a full read, as they
+% verse on ``(Artificial) Life and Dynamical Theory'' and on ``Body and Soul''
+% (``Non-Cartesian Robots'').
+
+Lemmen~\cite{lemmen1997} developed what he called a ``Spinozist'' framework for
+understanding cognition. Rooted on the work of Merleau-Ponty, one of its main
+points is the proposal of leaving Cartesianism behind, stating that
+
+\begin{quotation}
+  For the Spinozist, mind and body form a concrete unity.
+\end{quotation}
 

### 2011-12-21 draft.tex
Wrote on Van Gelder coupling, Freeman attractors and the relation to Merleau-Ponty.

@@
 \emph{intentional arc} and \emph{maximal grip}, respectively.
+
+% 2.C. Merleau-Ponty vs. Coupling (van Gelder) vs. Stepp vs. Freeman.
+% (Source: Thesis journal, pages 124 to 129, 11/03/2011).
+% (Source: Thesis journal, pages 90 to 99, 18/02/2011).
+% Note: Van Gelder ``coupling'' seems to be directly related to Freeman's basins
+% of attraction. I don't see how Stepp would come between them. 2.D is supposed
+% to be EST vs. discrete basin transitions, so it might belong down there. Maybe
+% we could cite our work on EST that also includes coupling.
+
+Van Gelder~\cite{vangelder1998} defends the \emph{dynamical hypothesis} in
+cognitive science, stating that
+
+\begin{quotation}
+  \dots cognitive agents ``are'' \dots not some particular dynamical systems but
+  as many systems as are needed to produce \dots cognitive performances \dots
+\end{quotation}
+
+He proposes that a cognitive system interacts with the environment through
+\emph{coupling}, with input influencing the direction of change continuously and
+output doing the same to the external world. Two different systems forging each
+other's transformations at the same time.
+
+Freeman~\cite{freeman1991}, after a set of experiments with rabbits, developed
+an \emph{attractor theory}, where the brain can be understood as a dynamical 
+system. According to this theory, the brain is composed of a set of energy peaks
+and valleys (formed from one's past experiences). When the subject is in a
+specific type of situation, the corresponding part of the brain landscape is
+activated, and movements are caused to get the brain closer to a state of
+minimum energy, called a ``basin of attraction''.
+
+Van Gelder's idea of a cognitive system being direcly connected to the world can
+be seen as an instance of Merleau-Ponty's \emph{intentional arc}, where
+Freeman's results support the validity of the \emph{maximal grip} concept.
 

### 2011-12-21 draft.tex
Wrote on the relation of the work to affordances.

@@
 \end{quotation}
+
+% 1.C. Relation to affordances.
+% (Source: Thesis journal, pages 130 to 139, 16/03/2011).
+% Note: the following two paragraphs only introduce affordances and a reason
+% for using it in robotics. Citing Chemero (2003) might be needed too. Examples
+% of affordance based robotics (from Rome's book) should be in section 3.
+
+Gibson~\cite{gibson1979} defines the \emph{affordances} of the environment as
+
+\begin{quotation}
+  \dots what it \emph{offers} the animal, what it \emph{provides} or
+  \emph{furnishes}, either for good or ill. \dots [S]omething that refers to
+  both the environment and the animal \dots.
+\end{quotation}
+
+Rome, Hertzberg and Dorffner~\cite{rome2008}, in their preface to ``Towards
+Affordance-Based Robot Control'', affirm that the concept of affordance has
+rarely been used in robotics or artificial intelligence, despite the original
+perspective it offers on coupling perception, action and reasoning.
 

### 2011-12-22 draft.tex
Started writing about affordances in robotics.

@@
 
+% 3.B. Affordances vs. Robotics (Rome's book).
+% (Source: Thesis journal, pages 130 to 139, 16/03/2011).
+% Note: I believe citing examples of affordances in Robotics is a must here.
+% A bridge between the former subject (Todes-Kuipers-Lemmen) might be necessary,
+% or even a new section.
+
+% Note: the next reference being focused on the coupling between more than one
+% robot, I'm not sure it fits completely here (in affordances) instead of
+% somewhere else (in coupling) or nowhere at all.
+Hafner and Kaplan~\cite{hafner2008} present the concept of \emph{interpersonal
+maps}, creating a representation space that allows one's behavior to the
+behavior of others. To validate their approach, they perform a set of
+experiments with four-legged robots either acting alone or engaged in imitation.
+
+Hertzberg et al.~\cite{hertzberg2008} discuss the perception of
+\emph{navegability} as an affordance. He presents \emph{navegability maps} as an
+example, and classify them as \emph{semantic maps}. They defend that
+
+\begin{quotation}
+  [E]very efficient robot control system includes highly tuned sensor data
+  processing modules \dots to perceive \dots robot-environment relations that
+  are necessary for the robot's intended function.
+\end{quotation}
+
+They leave, however, the insight to develop each of this processing modules to
+the designer, making it impractical for robots that should work in unstructured,
+dynamic environments.
+
+Raubal and Maratz~\cite{raubal2008} propose a functional model based on an
+extended theory of affordances. They demonstrate its applicability with two
+different robot scenarios, where the objects ``tell'' the robot what they can be
+used for.
+
+Sutton and Start~\cite{sutton2008} implement a system that segments images in
+\emph{affordances} instead of \emph{objects}. The training dataset is labeled is
+labeled manually. Even though, their algorithm is able to evaluate if a surface
+``provides stable support'' (like a chair), ``confirms graspability'' (like any
+graspable object) or ``confirms containment'' (like a cup). One of their results
+worth mentioning is the labelling of an upside-down garbage can as ``sittable''.
+
 % 4. The IKEA scenario.

### 2011-12-22 draft.tex
Finished writing about affordances in robotics.

@@
 
+Moratz and Tenbrink~\cite{moratz2008} discuss advances in \emph{spatial
+cognition} and develop a method that allows a robot to acquire knowledge through
+its interaction with a human. They also present an object recognition system
+based in \emph{affordances}, but do not provide detailed on the matter.
+
+Paletta and Fritz~\cite{paletta2008} describe a process of \emph{affordance
+cueing} where a robot obtains a probability distribution over the possible
+affordances from a set of features. They apply it to a scenario where a robot
+interacts with magnetizable cans that might afford ``lifting'', using
+Reinforcement Learning and Markov Decision Process to learn which features
+might indicate the ``liftability'' of an object.
+
 % 4. The IKEA scenario.

### 2011-12-26 draft.tex
Added EST, EST-Paladyn, Chemero. Compared everyone with Merleau-Ponty, Van Gelder and Freeman.

@@
 
+% 2.D. EST vs. Freeman (discrete basin transitions).
+% (Source: our BICA10 paper).
+% (Source: our Paladyn paper).
+% Note: I didn't make the comparison with the discrete basin transitions from
+% Freeman here because I believe it would belong together (and better) in the
+% end of the section.
+
+Event Segmentation Theory provides a model of how the human brain segments
+perception into a sequence of events~\cite{kurby07,zacks07}. This model sustains
+that event segmentation is based on the detection of prediction errors in the
+sensory stream. In particular, the human brain is permanently making predictions
+and comparing them with the actual outcome. Events are detected whenever a
+significant disparity between prediction and outcome is encountered.
+
+Nery and Ventura~\cite{nery2011} address the problem of segmenting a robot's
+perception into meaningful events, based on Event Segmentation Theory. In this
+approach, the prediction of the environment evolution is done using a dynamical
+systems approach, where the predictor system has a delayed coupling with the
+environment.
+
+% 2.E. Merleau-Ponty (demand) vs. Chemero
+% (Source: Thesis journal, pages 100 to 109, 03/03/2011).
+
+Chemero~\cite{chemero2003} criticizes Gibson's original definition of
+affordances, deeming it as deceitful. He defends that the perception of the
+environment is always connected to the animal, calling the perceived items
+\emph{features of situations}. He points out that are not directly related to
+``body scale'', but rather to the animal's ``abilities'', arguing that these are
+functional properties of the animal that happened to be historically useful.
+
+Chemero defines organisms as ``ability sets'' (and some abilities depend on
+others to function, e.g., walking depends on standing). Which abilities can be
+used depend on the current situation. He defends that an ``affordance layout''
+exists, and that this layout is changed by ``events''.
+
+% Note: The paragraph below tries to put together everything that was described
+% above: Van Gelder's coupling is the intentional arc; Freeman's basis of
+% attraction implement the maximal grip. Our work connects events and coupling,
+% justifying the corroboration of events to Chemero's affordance layout and its
+% changes caused by ``events''. Finally, Chemero's concept of affordances itself
+% can be seen as an instance of the intentional arc (opportunities presented by
+% the environment that ``call us'' to act on it), but on a different timescale.
+
 Van Gelder's idea of a cognitive system being direcly connected to the world can
@@
 Freeman's results support the validity of the \emph{maximal grip} concept.
+The work of Nery and Ventura shows the connection between the segmentation of
+perception into events and the eventual coupling of an organism and its
+environment. Chemero's idea of events changing the layout of affordances is
+corroborated by Event Segmentation Theory. The way Chemero presents the
+concept of affordances reinforces the idea of \emph{intentional arc}, closing
+the loop.
 
